# EDA 3 - Spatial Data

## Joins and sf objects

Spatial data typically consists of 2 parts: a normal old data frame and a geometry for each observation. These two parts are often contained in separate data files. So, the first topic of this section will be joining together different sets of data. We will look at the spatial distribution of crime in the US by joining together data from the <code>USArrests</code> dataset with data from the <code>us_states</code> dataset, which is contained in the package <code>spData</code>.

The object we will use to contain variable data as well as geographic data is an <code>sf</code> object, from a package with the same name. 

```{r echo=TRUE}
USArrests$name = rownames(USArrests) # add the state names as a variable
joined_df = left_join(USArrests, us_states, by = c("name" = "NAME")) # join the arrests and states geo data
joined_df_sf = st_as_sf(joined_df) # coerce to an sf object
```

Now, we have an <code>sf</code> object that we can use for plotting.

## Plotting spatial data 

To plot the number of murders by state, we can use ggplot with the geom function, <code>geom_sf</code>.

```{r echo=TRUE}
ggplot(joined_df_sf) + geom_sf(aes(fill = Assault))
```

Another useful 

```{r echo=TRUE}
mapview(joined_df_sf["Murder"])
```

## Assignment 3

To do this assignment, install the package <code>spData</code>, which contains the data you will use.

1. Join the <code>worldbank_df</code> with the <code>world</code> sf. Take a look at what each data set contains and what is a good variable to join on. Using that variable, join the two using a <code>left_join</code>. 

2. With your combined data set (an sf), visualize the population growth in African countries.

3. With your combined data set (an sf), visualize the urban population as a proportion of total population of European countries.

4. Is there are an association between life expectancy and population growth?


## ACS Data

We will look at American Community Survey (ACS) data to explore time and space concepts. What is the ACS:

- Annual survey of 3.5 million US households
- Covers topics not available in decennial US Census data (e.g. income, education, language, housing characteristics)
- Available as 1-year estimates (for geographies of population 65,000 and greater) and 5-year estimates (for geographies down to the block group)
- Data delivered as estimates characterized by margins of error

We will use a package called <code>tidycensus</code> to explore ACS data. To use tidycensus, you will need a Census API key. Visit [this page](https://api.census.gov/data/key_signup.html) to request a key, then activate the key from the link in your email.

Once activated, use the <code>census_api_key()</code> function to set your key as an environment variable.

```{r echo=TRUE, eval=FALSE}
library(tidycensus)
census_api_key("YOUR KEY GOES HERE", install = TRUE)
```

By default, ACS data are returned by <code>get_acs()</code> with the following five columns:

<code>GEOID</code>: unique identifier for the Census geographic unit

<code>NAME</code>: A descriptive name for the geographic unit (e.g. a state name)

variable: the ACS variable ID

<code>estimate</code>: The ACS estimate. Estimates are interpretated as characteristics rather than true counts.

<code>moe</code>: The margin of error associated with the estimate at a 90 percent confidence level.

Using the argument <code>geometry = T</code> will include a multipolygon for each row. This way, the data is an <code>sf</code> object on which you can use the plotting tools above. Here are some examples

### Median age by state

```{r echo=TRUE, warning=FALSE, results = 'hide', eval=FALSE}
median_age = get_acs(
  geography = "state",
  variables = "B01002_001",
  year = 2021,
  survey = "acs1",
  geometry = T
)
```

### Median age within Strafford County

```{r echo=TRUE, warning=FALSE, results = 'hide', eval=FALSE}
strafford_median_age <- get_acs(
  geography = "tract",
  variables = "B01002_001",
  state = "NH",
  county = "Strafford",
  year = 2021,
  geometry = T
)
```

### Median home value by county

```{r echo=TRUE, warning=FALSE, results = 'hide', eval=FALSE}
median_home_value <- get_acs(
  geography = "county",
  variables = "B25077_001",
  survey = "acs1",
  year = 2019,
  geometry = T
)
```

## Assignment 4

Choose a variable in the ACS1 survey that you think is interesting. See how it changes from 2005 - 2021 (by year)? Did the change(s) depend on geography? Make the best possible visualization you can for this.

