--- 
title: "Artifex - Spring 2023"
author: "David Reynolds"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
# output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: openscapes/series
description: "(Partial) course materials for Artifex, Fall 2022."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(broom)
library(lubridate)
library(knitr)
library(forecast)
library(nycflights13)
library(spData)
library(sf)
library(mapview)
```

# Course Overview - Artifex Spring 2023 {-}

Welcome (back) to Artifex! The primary objective of this course/ club is to equip its members with the essential skills of a data scientist. This is primarily achieved by working through a real world business analytics project with an industry partner. The course delivery is a mix of lectures and project-based learning. The lectures will all be tailored to our project and will be more prevalent in the beginning of the course. All lecture materials will be posted on this website. Further details on the partner project will be posted on Canvas. This class also has a significant club component, which mostly means that there is an emphasis on collaboration and getting to know one another.

## Prerequisites and Preparations {-}

- You should have some **basic knowledge** of R, and be familiar with the topics covered in the [Chapters 1 and 2 here](https://cengel.github.io/R-intro/).
- Have a recent version of [R](https://cran.r-project.org/) and [RStudio](https://www.rstudio.com/) installed. 
- Install and load the `tidyverse` package.

```{r, eval=FALSE, echo=TRUE}
install.packages("tidyverse")  
library(tidyverse)
```


## References {-}

Grolemund, G & Wickham, H (2017): R for Data Science http://r4ds.had.co.nz

Hyndman, R & Athanasopoulos, G (2022): Forecasting: Principles and Practice https://otexts.com/fpp2/index.html

