# Data Manipulation

We will discuss data transformation using data for all flights that departed NYC (i.e. JFK, LGA or EWR) in 2013. This data is contained in the <code>flights</code> data frame that comes from the <code>nycflights13</code> package. Before beginning, install this package and load it (using <code>nycflights13</code>).

There are five key dplyr functions that allow you to solve the vast majority of your data manipulation challenges:

1. Pick observations by their values (<code>filter()</code>).
2. Reorder the rows (<code>arrange()</code>).
3. Pick variables by their names (<code>select()</code>).
4. Create new variables with functions of existing variables (<code>mutate()</code>).
5. Collapse many values down to a single summary (<code>summarise()</code>). This is often used in conjunction with <code>group_by()</code>, which changes the scope of each function from operating on the entire dataset to operating on it group-by-group.

Let's go through a quick use case for each of these. Further detail can be found [here](https://r4ds.had.co.nz/transform.html).

## Filter

Suppose we want to look at just morning flights (before 10am).

```{r echo=TRUE, eval=FALSE}
morning_df = filter(flights, hour < 10)
```

How about morning flights in January. 

```{r echo=TRUE, eval=FALSE}
morning_df = filter(flights, hour < 10, month == 1)
```

## Arrange

What are the attributes of the flights with the biggest departure delays?

```{r echo=TRUE, eval=FALSE}
arrange(flights, desc(dep_delay))
```

## Select

Suppose we don't care about the flight number. 

```{r echo=TRUE, eval=FALSE}
arrange(flights, !flight)
```

## Mutate

Suppose we want to add a speed variable.

```{r echo=TRUE, eval=FALSE}
mutate(flights, speed = distance/air_time * 60)
```

## Group-By/ Summarise

Suppose we want to know the mean departure delay by airline.

```{r echo=TRUE, eval=FALSE}
flights %>% 
  group_by(carrier) %>% 
  summarize(mean_delay = mean(dep_delay, na.rm = T), n = n()) %>% 
  arrange(desc(mean_delay))
```


## Assignment 2

1. How many flights ... 

a. Had an arrival delay of two or more hours
b. Flew to Houston (IAH or HOU)
c. Were operated by United, American, or Delta
d. Departed in summer (July, August, and September)
e. Arrived more than two hours late, but didn’t leave late
f. Were delayed by at least an hour, but made up over 30 minutes in flight

2. Add two new variables to flights that convert <code>dep_time</code> and <code>arr_time</code> to a more convenient representation of number of minutes since midnight.

3. Look at the number of canceled flights per day. Is there a pattern over the course of the year? How about the proportion of canceled flights over the course of the year?

4. What time of day should you fly if you want to avoid delays as much as possible?





