# Statistical Analysis

## Time series

A central concept in time series analysis is autocorrelation. This is the correlation between $y_t$ and its lagged value. For a lag of $h$, this is the correlation between $y_t$ and $y_{t-h}$. Let's first review correlation. For two vectors of data, $x$ and $y$, the correlation between the two is,

\begin{align}
\text{cor}(x,y) &= \frac{\sum_i (x_i - \bar{x})(y_i - \bar{y})}{\sigma_x \sigma_y}
\end{align}

```{r echo=TRUE, eval=FALSE}
set.seed(1)
n = 5
x = rnorm(n)
y = rnorm(n)

# correlation
sum((x - mean(x)) * (y - mean(y)) ) / ((n-1)*sd(x) * sd(y)) # 0.2789049
cor(x, y) # 0.2789049
```


The autocorrelation takes this concept to a single time series. Autocorrelation, sometimes known as serial correlation in the discrete time case, is the correlation of a signal with a delayed copy of itself as a function of delay. The autocorrelation for a time series $y$ at lag $k$ is:

\begin{equation}
r_k = \frac{ \sum_{t = k + 1}^{T} (y_t - \bar{y})(y_{t-k} - \bar{y}) }{\sum_{t=1}^T (y_t - \bar{y})^2 }
\end{equation}

Here is a simple example of computing a lag 1 autocorrelation.

```{r echo=TRUE, eval=FALSE}
a = c(1,2,3,4,5)

a1 = c(1,2,3,4)
a2 = c(2,3,4,5)

# lag 1 autocorrelation
sum( (a1 - mean(a)) * (a2 - mean(a))) / (sum( (a - mean(a))^2 ) )  # by hand
(acf(a))
```

This statistic can be found using the <code>acf</code> function. Let's use the <code>quandl</code> package to look at the acf of Google's stock. 

```{r echo=TRUE}
goog = getSymbols('GOOG', from='2020-10-07', to='2023-02-10',auto.assign = FALSE)
goog$diff = diff(goog$GOOG.Close)
plot(goog$diff)
acf(goog$diff[-1])
```

## Exercise

If your team's data has a time component(s), explore the ACF of all time series variables as well as their differenced values. What does this tell you? 

If your team doesn't have time indexed data, find a different security from quandl and explore its ACF. How would you explain this? Also, look at the differenced value of the security's close price and its ACF. Compute the lag 10 autocorrelation "by hand" and check using <code>acf</code>